Compare scANVI and SingleR annotations

### 2025-08-14 analyses/cell-type-neuroblastoma-04/exploratory-notebooks/scanvi-singler-comparison.Rmd
update posterior section and land at 0.75

# palette files
recoded_palette_file <- file.path(
  module_dir,
  "palettes",
  "harmonized-cell-type-palette.tsv"
)
```

@@

# set up palettes
# recoded to shared colors
recoded_palette_df <- readr::read_tsv(recoded_palette_file)
recoded_celltype_pal <- recoded_palette_df$hex_color
names(recoded_celltype_pal) <- recoded_palette_df$harmonized_label
```

@@

First, what is the distribution of posterior probabilities for assigned labels?
We'll look at this on a per-library basis to see if any libraries may have less confident annotations in general.

```{r fig.height=12, fig.width=15}
scanvi_df |>
  tidyr::separate(cell_id, into = c("library_id", "barcode"), remove = FALSE) |>
  dplyr::add_count(library_id) |>
  dplyr::mutate(library_id = glue::glue("{library_id} (N={n})")) |>
  ############ into ggplot ################
  ggplot() + 
  aes(x = scanvi_posterior) + 
  # ugly but _very visible_ color
  geom_histogram(bins = 100, fill = "magenta", color = "magenta") + 
  facet_wrap(vars(library_id), scales = "free_y", ncol = 7)
```

The posterior probabilities here are all _exceptionally_ high, consistently across all libraries, which tells us that on the whole, `scANVI` is confident in assignments.

Let's instead look at these probabilities on a per-cell type basis:

```{r fig.height=8, fig.width=12}
ggplot(scanvi_df) + 
  aes(x = scanvi_posterior) + 
  # ugly but _very visible_ color
  geom_histogram(bins = 100, fill = "magenta", color = "magenta") + 
  facet_wrap(vars(scanvi), scales = "free_y", ncol = 5) 
```

While annotation reliability is about the same on a per-library basis, it seems that certain cell types may be much easier to classify than others which may be a factor of how well represented they are in the `NBAtlas` reference.
We'll look at the median value for each as well as the number of cells that were annotated for each label:

```{r}
scanvi_df |>
  dplyr::group_by(scanvi) |>
  dplyr::summarize(
    median_pp = median(scanvi_posterior), 
    total_cells = dplyr::n()
  ) |>
  dplyr::arrange(median_pp)
```

There are very few cells in general associated with the cell types that have very low median posterior probabilities (`NKT cell` and `Migratory cDC`), which is comforting.

A `0.9` threshold for the posterior probability here seems reasonable given how high these values are in general. 
We'll therefore recode cells with posterior probabilities below 0.9 to `unknown_scanvi`.

```{r}
pp_threshold <- 0.9
scanvi_df <- scanvi_df |>
  dplyr::mutate(scanvi = ifelse(
    scanvi_posterior < pp_threshold, 
    "unknown_scanvi",
    scanvi
  ))
```

@@

We'll create a single data frame with UMAP coordinates and all cell types (consensus, `SingleR`, and `scANVI`):

>>>>

# palette files
recoded_palette_file <- file.path(
  module_dir,
  "palettes",
  "harmonized-cell-type-palette.tsv"
)
nbatlas_palette_file <- file.path(
  module_dir,
  "palettes",
  "nbatlas-plus-cell-type-palette.tsv"
)
```

@@

# set up palettes
recoded_palette_df <- readr::read_tsv(recoded_palette_file)
recoded_celltype_pal <- recoded_palette_df$hex_color
names(recoded_celltype_pal) <- recoded_palette_df$harmonized_label

nbatlas_palette_df <- readr::read_tsv(nbatlas_palette_file)
nbatlas_celltype_pal <- nbatlas_palette_df$hex_color
names(nbatlas_celltype_pal) <- nbatlas_palette_df$NBAtlas_label
```

@@

First, what is the distribution of posterior probabilities for assigned labels?
We'll look at the minimum, maximum, and median posterior probabilities across libraries in a table.

```{r}
scanvi_df |>
  tidyr::separate(cell_id, into = c("library_id", "barcode"), remove = FALSE) |>
  dplyr::group_by(library_id) |>
  dplyr::summarize(
    library_size = dplyr::n(), 
    min_posterior = min(scanvi_posterior), 
    median_posterior = median(scanvi_posterior), 
    max_posterior = max(scanvi_posterior)
  ) |>
  dplyr::arrange(library_size)
```

Across libraries, the median posterior probability is always exceptionally high at >=0.99, which suggests `scANVI` confidence is roughly the same across libraries.

Let's instead look at these probabilities on a per-cell type basis, this time with a ridge plot including a line for the median value for each distribution:

```{r fig.height=8, fig.width=8}
scanvi_df |>
  dplyr::group_by(scanvi) |>
  dplyr::mutate(scanvi = glue::glue("{scanvi} (n={dplyr::n()})")) |>
  ########### into ggplot ############
  ggplot() + 
    aes(y = forcats::fct_infreq(scanvi), x = scanvi_posterior) + 
    # show the median
    ggridges::stat_density_ridges(quantile_lines = TRUE, quantiles = 2, alpha = 0.7) +
    labs(x = "posterior probability", y = "scANVI label (number of cells)") 
```

While annotation reliability is about the same on a per-library basis, it seems that certain cell types may be much easier to classify than others which may be a factor of how well represented they are in the `NBAtlas` reference.
There are very few cells in general associated with the cell types that have very low median posterior probabilities (`NKT cell`, `Migratory cDC`, `Circulating NK cell`).

A `0.75` threshold seems like it may be a good middle-ground for preserving labels while removing low-confidence annotations. 
What effect will this have?

```{r}
sum(scanvi_df$scanvi_posterior < 0.75) / nrow(scanvi_df)
```

This will throw out a bit under 2% of cells:

```{r}
scanvi_df |>
  dplyr::filter(scanvi_posterior < 0.75) |>
  dplyr::count(scanvi) |>
  dplyr::arrange(desc(n))
```

We'll therefore recode cells with posterior probabilities below 0.75 to `unknown_scanvi`.

```{r}
pp_threshold <- 0.75
scanvi_df <- scanvi_df |>
  dplyr::mutate(scanvi = ifelse(
    scanvi_posterior < pp_threshold, 
    "unknown_scanvi",
    scanvi
  ))
```

@@

We'll create a single data frame with UMAP coordinates and all cell types (consensus, `SingleR`, and `scANVI`).
